---
title: "Correlation Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Correlation Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
fig.width=8, 
fig.height=5,
warning = FALSE, 
message = FALSE,
collapse = TRUE,
comment = "#>"
)
```

Correlation analysis (along with seasonal analysis) is one of the main elements of the descriptive analysis of time series data. Furthermore, those patterns and insights revealed in the descriptive analysis play a pivotal role in the building process of a forecasting model. A common example is the setting of the AR and MA processes of the ARIMA models, which can be done with the autocorrelation and partial autocorrelation functions. Typically, time-series data will have some degree of correlation with its past lags. This can be seen in the hourly temperatures, as the current temperature is most likely to be close to the one during the previous hour. Moreover, correlation is also could reveal seasonal patterns as it is most likely that the series will have a high correlation with its seasonal lags. The **TSstudio** provides several tools for correlation analysis, such as `ts_cor`, `ts_lags`, and `ccf_plot`.

## The ACF and PACF functions

Traditionally, the `acf` (autocorrelation) and `pacf` (partial autocorrelations) functions from the **stats** package are used to calculate and plot the correlation relationship between the series and its lags. The `ts_cor` function provides an interactive wrapper and flexible version for those functions. Let's load the `UKgrid` series from the **UKgrid** package. This series represents the demand for electricity in the UK. Will use the `extract_grid` function to aggregate the series to daily frequency:


```{r }
library(UKgrid)

UKgrid_daily <- extract_grid(type = "ts",
             aggregate = "daily")

```

Before plotting the ACF and PACF of the series, let's review the series main characteristics:

```{r}
library(TSstudio)

ts_info(UKgrid_daily)

ts_plot(UKgrid_daily)
```


It is easy to observe, just by looking at the plot above, that the series has a strong seasonal component across the day of the year. If you zoom in, you can also observe that the series has a strong seasonal component across the day of the week (e.g., high consumption during the normal working days of the weeks and relatively low consumption throughout the weekend days). Quantify those relationships can be done by measuring the level of correlation between the series and its different seasonal lags. One of the downsides of the acf and pacf functions that it is hard to isolate a specific seasonal lag and compare that relationship over time. For example, the below plots represent the ACF and PACF relationship of the demand for electricity with its past 730 lags (or two seasonal cycles):

```{r fig.height=8}
par(mfrow=c(2,1))

acf(UKgrid_daily, 
    lag.max = 365 * 2)

pacf(UKgrid_daily, 
     lag.max = 365 * 2)
```

You can observe from the ACF plot above that the series is highly correlated with both its default seasonal lags (i.e., frequency 365) and also with the close lags. Yet, it is hard to compare between the two due to the high density of the plot. The ts_cor solve this issue by providing a more friendly plot of the ACF and PACF functions:

```{r, fig.height=8, fig.width=10}
ts_cor(UKgrid_daily, 
       seasonal = TRUE, 
       lag.max = 365 * 2)
```

The `seasonal` argument, when set to TRUE, allows you to distinguish the seasonal lags from the non-seasonal lags. In addition, you can remove the non-seasonal lags by clicking on the non-seasonal legend. This is very useful to tune the seasonal components of SARIMA model. The `seasonal_lags` argument allows you to mark and distinguish other seasonal lags, besides the default one. For instance, add to the plot the weekly lags by setting the `seasonal_lags` to 7:

```{r, fig.height=8, fig.width=10}
ts_cor(UKgrid_daily, 
       seasonal = TRUE, 
       lag.max = 365 * 2, 
       seasonal_lags = 7)
```

### Lags plot

Lags plot is alternative approach for visualize the relationship between the series and its lags. As linear the relationship between the series and its lags the highes the correlation between the two. For example, let's plot the relationship of the demand for electricity with its first 7 lags:

```{r}
ts_lags(UKgrid_daily, lags = c(1:7))

```

As observed with the ACF plot, the series has strong linear relationship with first and seventh lags. Similarly, we can plot the relationship of the series with its yearly lags and compared it with the seventh lag:

```{r}
ts_lags(UKgrid_daily, lags = c(7, 364 ,365))

```

```{r}
data("USUnRate")
data("USVSales")

ccf_plot(x = USVSales, y = USUnRate)

#Plotting the first 6 lead and lags of the USVSales with the USUnRate
ccf_plot(x = USVSales, y = USUnRate, lags = -12:12)

```

