---
title: "Forecasting Applications"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Forecasting Applications}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
fig.width=8, 
fig.height=5,
warning = FALSE, 
message = FALSE,
collapse = TRUE,
comment = "#>"
)
```

The **TSstudio** package provides a set of functions for train, test, and evaluate time series forecasting models, this vignette covers the main ones.


## Creating testing and training partitons

The `ts_split` function split time series data into training (sample-in) and testing (sample-out) partitions, keeping the chronological order of the series. The `sample.out` argument defines the length of the testing partition. In the following example, we will use the function to split the `USgas` series (the monthly demand for natural gas in the US) into training and testing partitions, leaving the last 12 months as testing partition:

```{r}
library(TSstudio)

data(USgas)

ts_info(USgas)

USgas_par <- ts_split(USgas, sample.out = 12)

train <- USgas_par$train

test <- USgas_par$test

ts_info(train)
ts_info(test)


```


## Forecast perfromance 

The `test_forecast` function visualizes the performance of a forecasting model on the training and testing partitions by plotting the fitted and forecasted values against the actuals. The function supports forecast models from the `forecast` and `stats` packages such as `auto.arima`, `tslm`, `ets`, `HoltWinters`, etc.


For example, let's train a `tslm` model on the `train` partition and forecast the corresponding observations of the `test` partition:

```{r}
library(forecast)

md <- tslm(train ~ season + trend)

fc <- forecast(md, h = 12)
```

Using the actual series (`USgas`), forecast object (`fc`), and the corresponding testing partition (`test`), the function will plot the fitted and forecasted values against the actuals:

```{r}
test_forecast(actual = USgas,
              forecast.obj = fc,
              test = test)
```

The MAPE and RMSE score of both the fitted and forecasted values can be seen when hovering on the plot. 

# Training multiple time series model with backtesting

The `train_model` function provides a flexible framework for forecasting. The use of the function is for train, test, and evaluate multiple models (or different versions of those models) with the use of a backtesting approach. The backtesting approach enables you to evaluate the performance of the model over time (as opposed to single training and testing partitions) and select a model that performed best using some error criteria (e.g., **MAPE** or **RMSE**). The main arguments of this function are:


`methods` - defines the type of models and their setting to be used on the training process. Currently, it supports the following models:

  + `arima` - model from the stats package

  + `auto.arima` - model from the forecast package

  + `ets` - model from the forecast package

  + `HoltWinters` - model from the stats package

  + `nnetar` - model from the forecast package

  + `tslm` - model from the forecast package (note that the 'tslm' model must have the formula argument in the 'method_arg' argument)
  

```{r}
methods <- list(ets1 = list(method = "ets",
                            method_arg = list(opt.crit = "lik"),
                            notes = "ETS model with opt.crit = lik"),
                ets2 = list(method = "ets",
                            method_arg = list(opt.crit = "amse"),
                            notes = "ETS model with opt.crit = amse"),
                arima1 = list(method = "arima",
                              method_arg = list(order = c(2,1,0)),
                              notes = "ARIMA(2,1,0)"),
                arima2 = list(method = "arima",
                              method_arg = list(order = c(2,1,2),
                                                seasonal = list(order = c(1,1,1))),
                              notes = "SARIMA(2,1,2)(1,1,1)"),
                hw = list(method = "HoltWinters",
                          method_arg = NULL,
                          notes = "HoltWinters Model"),
                tslm = list(method = "tslm",
                            method_arg = list(formula = input ~ trend + season),
                            notes = "tslm model with trend and seasonal components"))
# Training the models with backtesting
md <- train_model(input = USgas,
                  methods = methods,
                  train_method = list(partitions = 4, 
                                      sample.out = 12, 
                                      space = 3),
                  horizon = 12,
                  error = "MAPE")

plot_model(md)

plot_error(md)
```

