---
title: "Forecasting Applications"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Forecasting Applications}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
fig.width=8, 
fig.height=5,
warning = FALSE, 
message = FALSE,
collapse = TRUE,
comment = "#>"
)
```

The **TSstudio** package provides a set of functions for train, test, and evaluate time series forecasting models. This vignette covers the main ones.


### Creating testing and training partitons

The `ts_split` function split time series data into training (sample-in) and testing (sample-out) partitions, keeping the chronological order of the series. The `sample.out` argument defines the length of the testing partition. In the following example, we will use the function to split the `USgas` series (the monthly demand for natural gas in the US) into training and testing partitions, leaving the last 12 months as testing partition:

```{r}
library(TSstudio)

data(USgas)

ts_info(USgas)

USgas_par <- ts_split(USgas, sample.out = 12)

train <- USgas_par$train
test <- USgas_par$test

ts_info(train)
ts_info(test)


```


```{r}
library(forecast)
md <- tslm(train ~ season + trend)
fc <- forecast(md, h = 12)

test_forecast(actual = USgas,
              forecast.obj = fc,
              test = test)
```




```{r}
 hw_grid_shallow <- ts_grid(ts.obj = USgas,
                            periods = 6,
                            model = "HoltWinters",
                            optim = "MAPE",
                            window_space = 6,
                            window_test = 12,
                            hyper_params = list(alpha = seq(0.01, 1,0.1),
                                                beta =  seq(0.01, 1,0.1),
                                                gamma = seq(0.01, 1,0.1)),
                            parallel = TRUE,
                            n.cores = 8)

TSstudio::plot_grid(hw_grid_shallow)

 a_min <- min(hw_grid_shallow$grid_df$alpha[1:20])
 a_max <- max(hw_grid_shallow$grid_df$alpha[1:20])
 
 b_min <- min(hw_grid_shallow$grid_df$beta[1:20])
 b_max <- max(hw_grid_shallow$grid_df$beta[1:20])
 
 g_min <- min(hw_grid_shallow$grid_df$gamma[1:20])
 g_max <- max(hw_grid_shallow$grid_df$gamma[1:20])
 
 hw_grid_second <- ts_grid(ts.obj = USgas,
                           periods = 6,
                           model = "HoltWinters",
                           optim = "MAPE",
                           window_space = 6,
                           window_test = 12,
                           hyper_params = list(alpha = seq(a_min, a_max,0.05),
                                               beta =  seq(b_min, b_max,0.05),
                                               gamma = seq(g_min, g_max,0.05)),
                           parallel = TRUE,
                           n.cores = 8)


TSstudio::plot_grid(hw_grid_second)
TSstudio::plot_grid(hw_grid_second, type = "3D")

md <- HoltWinters(USgas, 
                   alpha = hw_grid_second$alpha,
                   beta = hw_grid_second$beta,
                   gamma = hw_grid_second$gamma)
 
 
fc <- forecast(md, h = 60)
 
 plot_forecast(fc)
```

```{r}
methods <- list(ets1 = list(method = "ets",
                            method_arg = list(opt.crit = "lik"),
                            notes = "ETS model with opt.crit = lik"),
                ets2 = list(method = "ets",
                            method_arg = list(opt.crit = "amse"),
                            notes = "ETS model with opt.crit = amse"),
                arima1 = list(method = "arima",
                              method_arg = list(order = c(2,1,0)),
                              notes = "ARIMA(2,1,0)"),
                arima2 = list(method = "arima",
                              method_arg = list(order = c(2,1,2),
                                                seasonal = list(order = c(1,1,1))),
                              notes = "SARIMA(2,1,2)(1,1,1)"),
                hw = list(method = "HoltWinters",
                          method_arg = NULL,
                          notes = "HoltWinters Model"),
                tslm = list(method = "tslm",
                            method_arg = list(formula = input ~ trend + season),
                            notes = "tslm model with trend and seasonal components"))
# Training the models with backtesting
md <- train_model(input = USgas,
                  methods = methods,
                  train_method = list(partitions = 4, 
                                      sample.out = 12, 
                                      space = 3),
                  horizon = 12,
                  error = "MAPE")

plot_model(md)

plot_error(md)
```

