---
title: "Plotting Time Series Data"
author: "Rami Krispin (@Rami_Krispin)"
date: "`r Sys.Date()`"
output: html_document
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{Plotting Time Series Data}
  %\usepackage[UTF-8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(list(menu.graphics = FALSE, scipen=99, digits= 3))
```

The `ts_plot` is a versatile function for interactive plotting of time series objects. It is based on the [plotly](https://plot.ly/r/) engine and supports multiple time series objects such as `ts`, `mts`, `zoo`, `xts` and as well the data frame family (`data.frame`, `data.table`, and `tbl`).

```{r}
# install.packages("TSstudio")
library(TSstudio)
packageVersion("TSstudio")
# Function documentation
?ts_plot
```


#### Plotting a univariate time series

Plotting a univariate time series data with the `ts_plot` function is straightforward: 

```{r }
# Loading the USgas dataset
data("USgas")

ts_info(USgas)

```

```{r fig.height=5, fig.width=10}
ts_plot(USgas)
```



By default, the function will set the input object name as the plot title and leave the X and Y axises empty. It is possible to label the axises and set a different title:

```{r fig.height=5, fig.width=10}
# Setting the plot titles
ts_plot(USgas,
        title = "US Natural Gas Consumption",
        Xtitle = "Year",
        Ytitle = "Billion Cubic Feet")
```

#### Line setting

There are several arguments which allow modifying the line main characteristics such as the colors, width, and type of line. 

The `line.mode` argument is equivalent to the `mode` argument of the `plot_ly` function, and there are 3 options:

1. `line` - the default option, a clean line
2. `line+markers` - a line with a markers
3. `markers` - only markers

The `dash` argument allows modifying the line to dashed or dotted by setting the argument to `dash` or `dot` respectively.  The line width, by default, set to 2 and can be modified with the `width` argument:

```{r fig.height=5, fig.width=10}
ts_plot(USgas,
        title = "US Natural Gas Consumption",
        Xtitle = "Year",
        Ytitle = "Billion Cubic Feet",
        line.mode = "lines+markers",
        width = 3,
        color = "green")
```

In addition, we can add grid lines for the Y and X axises and slider by setting the `Xgrid`, `Ygrid` and slider arguments to TRUE:

```{r fig.height=5, fig.width=10}
ts_plot(USgas,
        title = "US Natural Gas Consumption",
        Xtitle = "Year",
        Ytitle = "Billion Cubic Feet",
        line.mode = "lines+markers",
        width = 1,
        color = "#66C2A5",
        Xgrid = TRUE,
        Ygrid = TRUE,
        slider = TRUE)
```

#### Ploting a multiple time series object

The `ts_plot` can handle multiple time series objects such as `mts`, `xts`, `zoo` and data frame family objects. The example below demonstrates the plot options for a multiple time series object with the closing prices of **Apple**, **Facebook**, **Google** and **Microsoft** stocks over time:
```{r fig.height=5, fig.width=10, message=FALSE, warning=FALSE}
library(TSstudio)
library(xts)
library(zoo)
library(quantmod)
# Loading the stock price of key technology companies:
tckrs <- c("GOOGL", "FB", "AAPL", "MSFT")
getSymbols(tckrs, 
           from = "2013-01-01",
           src = "yahoo"
           )

# Creating a multiple time series object
closing <- cbind(AAPL$AAPL.Close, FB$FB.Close, GOOGL$GOOGL.Close, MSFT$MSFT.Close)
names(closing) <- c("Apple", "Facebook", "Google", "Microsoft")

ts_info(closing)
```

The `type` argument defines whatever to plot all the series in a single plot (`single` option) or plot each series separately (`multiple` option):

```{r fig.height=5, fig.width=10}
# Plot each series in a sepreate (default option)
ts_plot(closing,
        title = "Top Technology Companies Stocks Prices Since 2013", 
        type = "multiple")


# All the series in one plot
ts_plot(closing, 
        title = "Top Technology Companies Stocks Prices Since 2013",
        type = "single")


```

#### Plotting data frame objects

Currently, the `ts_plot` function supports three classes of data frame - `data.frame`, `data.table` and `tbl`. To be able to plot a data frame object, it must contain one column with a date or time object (`Date` or `POSIXlt`/`POSIXct`) and at least one numeric column. The **US_indicators** is an example of a `data.frame` object with time series data. It contains the monthly vehicle sales and the unemployment rate in the US since 1976 and a date object:

```{r}
data(US_indicators)
str(US_indicators)
```

```{r fig.height=5, fig.width=10}
ts_plot(US_indicators)
```

