---
title: "Plotting Time Series Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plotting Time Series Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  fig.width=8, 
  fig.height=5,
  collapse = TRUE,
  comment = "#>"
)
```

## The ts_plot function

The plotting of time series object is most likely one of the steps of the analysis of time-series data. The \code{\link[TSstudio]{ts_plot} is a customized function for plotting time series data based on the [plotly](https://plot.ly/r/) package visualization engine. It supports the following time-series classes:

* ts
* mts
* zoo
* xts
* data.frame^1^ 
* tbl^1^ 

^1^ Must have a `Date` or `POSIXct/lt` column and at least on numeric column

For example let's plot the `USgas` series, a `ts` object:

```{r setup}
library(TSstudio)
```

```{r message=FALSE, warning=FALSE}
data(USgas)

ts_info(USgas)

ts_plot(USgas)
```

## Customization options

It is straightforward to customize the plot using the different arguments of the ts_plot function

### Adding titles

By default, the function sets the series name as the plot title. The `title` allows you to modify the plot title and the `Xtitle` and `Ytitle` enable you to add titles to the X-axis and Y-axis respectively:

```{r}
ts_plot(USgas,
        title = "US Monthly Natural Gas Consumption",
        Xtitle = "Time",
        Ytitle = "Billion Cubic Feet")
```

### Adding slider

The `slider` argument allows you to add a slider on the bottom of the plot, which will enable you to customize the window length of the plot:

```{r}
ts_plot(USgas,
        title = "US Monthly Natural Gas Consumption",
        Xtitle = "Time",
        Ytitle = "Billion Cubic Feet",
        slider = TRUE)
```


### Customize the line 


