---
title: "Seasonality Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Seasonality Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
fig.width=8, 
fig.height=5,
warning = FALSE, 
message = FALSE,
collapse = TRUE,
comment = "#>"
)
```


Seasonality analysis is one of the core elements of the descriptive analysis process of time series data. This process is typically done with the use of data visualization tools and summary statistics methods. The **TSstudio** package provides a set of functions for seasonality plots.

## The ts_decompose function

The ts_decompose function is an interactive wrapper for the `decompose` function from the **stats** package. It provides a decomposition of the series into seasonal, trend, and irregular components. The decomposition of the series is done by using a moving average.  The following example demonstrated the decomposition of the US monthly consumption of natural gas (`USgas`):

```{r}
library(TSstudio)

data(USgas)

ts_info(USgas)

ts_decompose(USgas)
```





## The ts_seasonal plot

The `ts_seasonal` function provides three different seasonal views of the series for low-frequency time series objects (i.e., monthly, quarterly, half-year). The type of seasonal view defined by the `type` argument. The default view, `type = normal`, split and plot the series by its full cycle units (which is typically years). This allows comparing the changes of the series from year to year by the series frequency units (e.g. the month of the year), and therefore, to identify seasonal patterns:

```{r}
ts_seasonal(USgas, type = "normal")
```

Note that the color scale of the series cycles lines (which in the case above represents the years) is set by chronological order. The second view, `type = cycle`, split and plot the series by its frequency units over time. This enables to identify if there is a hierarchy relationship between the frequency units of the series, or how strong the seasonal pattern of the series:

```{r}
ts_seasonal(USgas, type = "cycle")
```

In the plot above, you can notice that while the series is trending up, the hierarchy between the different months kept over time (e.g., typically, the peak of the consumption occurs during January). The third and last view, `type = box`, returns a box-plot of each frequency unit:

```{r}
ts_seasonal(USgas, type = "box")
```

As each view may reveal different patterns, it would make more sense to plot all the three together. This can be done by setting `type = all`:

```{r}
ts_seasonal(USgas, type = "all")
```

Note that some of the variations seen in those seasonal plots, especially in the `box` view,  related to the series trend. In some cases, it would make more sense to remove the trend from the series and replot it. The next example demonstrated the plot of the series without the trend. A simple detrending of the series is done with the decompose function:

```{r}
USgas_detrend <- USgas - decompose(USgas)$trend
ts_seasonal(USgas_detrend, type = "all")
```


## Time series heatmap

Another way to look at seasonality is with the use of heatmap. The ts_heatmap function returns heatmap of the series object:

```{r}
ts_heatmap(USgas)
```

The `color` argument defines the heatmap color scale supporting the RColorBrewer colors palettes. For example, we can modify the heatmap above to green scale by using the `Greens` colors palette:

```{r}
ts_heatmap(USgas, color = "Greens")
```


