---
title: "Introduction for the TSstudio Package"
author: "Rami Krispin"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## **Overview**

The TSstudio package provides a set of tools for time series analysis. That includes a variety of interactive visualizations tools, based on the [plotly](https://plot.ly/r/) package, for descriptive, seasonality and correlation analysis and as well forecasting supporting objects of the [forecast](https://github.com/robjhyndman/forecast) and other packages. 

## **Installation**

Install the stable version from [CRAN](https://CRAN.R-project.org/package=TSstudio):
```{r message=FALSE, warning=FALSE, eval=FALSE, echo=TRUE}
install.packages("TSstudio")
```


or install the development version from [Github](https://github.com/RamiKrispin/TSstudio):
```{r message=FALSE, warning=FALSE, eval=FALSE, echo=TRUE}
# install.packages("devtools")
devtools::install_github("RamiKrispin/TSstudio")
```


## **Key Features**

### Visualization of time series objects with the ts_plot function

The ts_plot function for visualization of time series objects support multiple types of time series class such as "ts", "mts", "xts", "zoo" and as well the data frame family ("data.frame", "tbl", "data.table") as long as it contains a date (Date or POSIXlt/ct) object.

```{r fig.height=5, fig.width=8, message=FALSE, warning=FALSE, include=TRUE}
library(TSstudio)
library(xts)
library(zoo)
library(quantmod)
# Loading the stock price of key technology companies:
tckrs <- c("GOOGL", "FB", "AAPL", "MSFT")
getSymbols(tckrs, 
           from = "2013-01-01",
           src = "yahoo"
           )

# Visual Google closing price since 2013
Google <- GOOGL$GOOGL.Close
class(Google)
head(Google)
# Basic plot
ts_plot(Google)


# Adding titles and slider
ts_plot(Google, 
        title = "Google Stock Prices Since 2013",
        Xtitle = "Sourch: Yahoo Finance", 
        Ytitle = "Closing Price in USD",
        slider = TRUE
        )


```

#### Plotting multiple time series data

It is possible to plot a multiple time series object ("mts", "xts", "zoo" or data frame), either in a single plot or on multiple plots (one for each): 

```{r fig.height=5, fig.width= 8, message=FALSE, warning=FALSE}
closing <- cbind(GOOGL$GOOGL.Close, FB$FB.Close, AAPL$AAPL.Close, MSFT$MSFT.Close)
names(closing) <- c("Google", "Facebook", "Apple", "Microsoft")

class(closing)
dim(closing)

# You can plot all the series in one plot using the type option:
ts_plot(closing, 
        title = "Top Technology Companies Stocks Prices Since 2013",
        type = "single")

# or use the default option - "multiple" and plot series on separate plots:
ts_plot(closing,
        title = "Top Technology Companies Stocks Prices Since 2013")

```

### Seasonality analysis

The TSstudio provides a variety of tools for seasonality analysis, currently, most of the functions support series with daily, weekly, monthly and quarterly frequency.

#### The ts_seasonal function

The ts_seasonal function provides the ability to plot a time series data by its frequency cycle (hence, for monthly frequency by a full year) or by frequency units (thus, for monthly data by the months of the year). For example, the USgas dataset, the US monthly consumption of natural gas (one of the package dataset), has a strong seasonal pattern:

```{r fig.height=5, fig.width= 8, message=FALSE, warning=FALSE}
# Load the US monthly natural gas consumption
data("USgas")

class(USgas)

ts_plot(USgas,
        title = "US Natural Gas Consumption",
        Xtitle = "Year",
        Ytitle = "Billion Cubic Feet"
        )

```

#### The ts_seasonal function

The ts_seasonal function provides 3 types of seasonal plots:

1. "normal" - type provides a break of the series by the cycle units (or years). This allows identifying whether the series has a reputable seasonal pattern: 

```{r fig.height=5, fig.width= 8, message=FALSE, warning=FALSE}
ts_seasonal(USgas, type = "normal")

```
The colors are being set automatically by a sequential color palette in chronical order.

2. "cycle" - this option provides a view of each frequency unit of the series across the full cycle units of the frequency, in the example below you can notice, for example, that the consumption during January in most of the years was the peak of the year:

```{r fig.height=5, fig.width= 8, message=FALSE, warning=FALSE}
ts_seasonal(USgas, type = "cycle")
```

3. "box" - for representing the cycle units with a box plot:

```{r fig.height=5, fig.width= 8, message=FALSE, warning=FALSE}
ts_seasonal(USgas, type = "box")
```


Alternatively, setting the type = “all”, print the three options above together in one plot:

```{r fig.height=5, fig.width= 8, message=FALSE, warning=FALSE}
ts_seasonal(USgas, type = "all")
```

Note: that the colors of the months in the "cycle" plot corresponding to the colors in the "box" plot. 

In addition, you can modify the color palette using one of the RColorBrewer or viridis palette options: 

```{r fig.height=5, fig.width= 8, message=FALSE, warning=FALSE}
ts_seasonal(USgas, type = "all", palette_normal = "inferno", palette = "Accent")
```



#### The ts_heatmap function

Another useful visualization tool for seasonality analysis is the ts_heatmap function for time series objects, where the y axis represents the cycle units and x axis represents the years:

```{r fig.height=5, fig.width= 8, message=FALSE, warning=FALSE}
ts_heatmap(USgas)
```

By default, the function is using the "Blues" palette of the RColorBrewer package. You can modify the palette by modifying the color argument:

```{r fig.height=5, fig.width= 8, message=FALSE, warning=FALSE}
ts_heatmap(USgas, color = "Reds")
```

One of the main improvement since the last release is the support for daily and weekly frequency in addition to the monthly and quarterly. For example, we will plot the daily demand for electricity in the UK (available in the UKgrid package):

```{r fig.height=5, fig.width= 8, message=FALSE, warning=FALSE}
#install_github("RamiKrispin/UKgrid")
library(UKgrid)

UKgrid_daily <- extract_grid(type = "tbl", aggregate = "daily")
head(UKgrid_daily)

ts_heatmap(UKgrid_daily, color = "BuPu")
```



Similarly, the heatmap could be represented by a surface 3d plot:  

```{r fig.height=5, fig.width= 8, message=FALSE, warning=FALSE}
ts_surface(USgas)
```

#### Polar plot

The ts_polar function provides a polar plot demonstrative of time series data where the year is represented by color and the magnitude is represented by the size of the cycle unit layer:
```{r fig.height=5, fig.width= 8, message=FALSE, warning=FALSE}
ts_polar(USgas)
```


#### Decomposing time series

The ts_decompose function transforms the decompose function output(of the stats package) to an interactive format. In addition to the "additive" and the "multiplicative" options of the original decompose function, there is the "both" option which represents both options side by side:

```{r fig.height=5, fig.width= 8, message=FALSE, warning=FALSE}
ts_decompose(USgas, type = "both")
```

### Correlation Analysis

Similarly to the decompose function, the TSstudio transforms the acf and pacf functions into a colorful version with the ts_acf and ts_pacf:

```{r, fig.height=5, fig.width= 8, message=FALSE, warning=FALSE}
ts_acf(USgas, lag.max = 36)
ts_pacf(USgas, lag.max = 36)

```

Another intuitive method to identify correlation between the series and its lags is with lag plot:

```{r fig.height=5, fig.width= 8, message=FALSE, warning=FALSE}
ts_lags(USgas)
```


### Applications for forecasting

#### Splitting time series object for training and testing partitions

The TSstudio package provides a set of supporting tools for training and testing forecasting models. The ts_split function splits the series into training and testing partitions based on a given horizon:

```{r fig.height=5, fig.width= 8, message=FALSE, warning=FALSE}
# set the forecast horizon for 12 months
h <- 12

# Split the data into training and testing sets (leaving the last 12 months for testing)
split_USgas <- ts_split(USgas, sample.out = h)

train <- split_USgas$train
test <- split_USgas$test


head(train, 5)
head(test, 5)

```

#### Residuals analysis

The check_res function visualize the traditional plot of the residuals over time, and the corresponding auto correlation function and the histogram (of the residuals):

```{r fig.height=5, fig.width= 8, message=FALSE, warning=FALSE}
library(forecast)
# Building a model on the training set
fit <- auto.arima(train, lambda = BoxCox.lambda(train))

# Checking the residuals
check_res(fit)

```

#### Forecast evaluation with the test_forecast

The test_forecast function visualizes the fitted values vs the training partition values and the forecasted values vs the testing partitions values. In addition, the tooltip of the plot provides information about the model performance (MAPE and RMSE of the training and testing partitions):

```{r fig.height=5, fig.width= 8, message=FALSE, warning=FALSE}
fc <- forecast(fit, h = h)

test_forecast(actual = USgas, forecast.obj = fc, test = test)

```

### Other utility tools

#### ts to data frame format

The ts_reshape function transforms time series data ("ts", "zoo" and "xts" classes) into a friendly data frame format using the years as the columns and the cycle units as rows (currently supporting only monthly and quarterly frequencies):

```{r fig.height=5, fig.width= 8, message=FALSE, warning=FALSE}
# Loading the Total US Vehicle Sales data 
data("USVSales")

ts_plot(USVSales, title = "Total US Vehicle Sales",
        Ytitle = "Thousands of Units",
        Xtitle = "Source: U.S. Bureau of Economic Analysis")

head(ts_plot)

USVSales_df <- ts_reshape(USVSales)

library(DT)

datatable(USVSales_df, filter = 'top', options = list(
  pageLength = nrow(USVSales_df), autoWidth = TRUE
))
```


#### Converting "zoo" or "xts" objects to "ts" class

The xts_to_ts or the zoo_to_ts functions convert "zoo" or "xts" objects into "ts" class:

```{r fig.height=5, fig.width= 8, message=FALSE, warning=FALSE}
# Loading the University of Michigan Consumer Sentiment 
data("Michigan_CS")

Michigan_CS_ts <- xts_to_ts(Michigan_CS)

ts_plot(Michigan_CS, title = "University of Michigan Consumer Sentiment - 'xts' format")
ts_plot(Michigan_CS_ts, title = "University of Michigan Consumer Sentiment - 'ts' format")

class(Michigan_CS)
class(Michigan_CS_ts)

head(Michigan_CS)
head(Michigan_CS_ts)

head(index(Michigan_CS))
head(time(Michigan_CS_ts))

periodicity(Michigan_CS)
frequency(Michigan_CS_ts)


cycle(Michigan_CS_ts)
```




