---
title: "Introduction for the TSstudio Package"
author: "Rami Krispin"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## **Overview**

The TSstudio (time series studio) package is a toolbox for interactive visualization applications for time series analysis with the [Plotly](https://plot.ly/r/) package. The package is under development and first iteration is available on [github](https://github.com/RamiKrispin/TSstudio).

## **Installation**

The package is available for installation with the devtools package (if devetools package is not installed please use `install.packages("devtools")` to install it).

```{r message=FALSE, warning=FALSE}
# Install the TSstudio
devtools::install_github("RamiKrispin/TSstudio")
```

## **Key Features**

### Visualization of fitted and forecasted values vs. actual values

```{r fig.height = 5, fig.width = 8}
library(TSstudio)
library(forecast)

# Load the AirPassengers dataset
data("AirPassengers")

# set the forecast horizon for 12 months
h <- 12 

# Split the data into training and testing sets (leaving the last 12 months for testing)
train <- window(AirPassengers, 
                start = time(AirPassengers)[1], 
                end = time(AirPassengers)[length(AirPassengers) - h])
test <- window(AirPassengers, 
               start = time(AirPassengers)[length(AirPassengers) - h + 1], 
               end = time(AirPassengers)[length(AirPassengers)])

# Using auto.arima to train and forecast the last 12 months on the series
fc <- forecast(auto.arima(train, lambda = BoxCox.lambda(train)), h = h)

# Plotting the series vs the fitted and the forecasted
fortest_ly(actual = AirPassengers, forecast.obj = fc, train = train, test = test)

```

### Seasonality plot

```{r fig.height = 5, fig.width = 8}
seasonal_ly(AirPassengers)
```


### Plotting time series object
```{r fig.height = 5, fig.width = 7}
# Simple plot
ts.plot_ly(AirPassengers)
```

```{r fig.height = 5, fig.width = 7}
# Adding slider and markers, changing the color to green
ts.plot_ly(AirPassengers, 
           slider = TRUE, 
           line.mode = "lines+markers", 
           color = "green", 
           width = 2)

```

### Plotting the acf and the pacf estimation

```{r fig.height=5, fig.width=8}
#plotting the acf and the pacf estimation
# setting the lags to 60 and the level of significant to 0.01
acf_ly(AirPassengers, lag.max = 60, ci = 0.99)
pacf_ly(AirPassengers, lag.max = 60, ci = 0.99)
```

